---
title: "Exploring Thompsons Creek Stage Discharge Data"
date: "`r Sys.Date()`"
github-repo: https://github.com/mps9506/thompson-stage-discharge
bibliography: bibliography.bib
biblio-style: "apalike"
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.retina = 2)
```

## About

This document is an exploratory analysis of the pressure data and IQ data at ...

```{r library, echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
library(tibble)
library(dplyr)
library(ggplot2)
library(stringr)
library(here)
library(units)
library(ggforce)
library(hrbrthemes)
library(lubridate)
library(purrr)
library(hydroGOF)


update_geom_font_defaults(font_rc)

theme_ms <- function(...) {
  theme_ipsum_pub(plot_margin = margin(10,10,10,10),
              axis_title_just = "c") +
    theme(legend.position = "bottom",
          panel.background = element_rect(fill = "white", 
            colour = NA), 
          panel.border = element_rect(fill = NA, 
            colour = "grey20"),
          ...)
}
```



## Data Import

### Pressure Transducers

Use `readr::read_csv()` to import data. The HOBO pressure data is pre-computed but has extra lines we need to clean up.

- each location has two files, March through August and August through December.

- ideally, I can make one dataframe that compiles all the data

- we can read all the files and row bind but we need to add a column indicating the site number

- for this deployment it looks like the daylight savings time adjustment is not applied to data. So everything between March 6 and November 1 is off 1 hour from DST. Assuming all data will be reported in observed time, it is probably easiest to convert time back to GMT, then apply R's built in time zones to convert the times with appropriate -5 or -6 hour offsets.

```{r hoboimport}

## make a list of files to import
file_paths <- paste0(here("Data/Hobo"),
                     "/",
                     list.files(path = here("Data/Hobo"),
                                pattern = ".csv"))

## create a blank tibble to fill
hobo_df <- tibble()


## loop through file paths to read each file
for (i in file_paths) {
  x <- read_csv(
    i,
    skip = 2,
    col_names = c(
      "Row",
      "Date",
      "Time",
      "Abs_Pres",
      "Temp",
      "Bar_Pressure",
      "Water_Level",
      "Coupler_Detached",
      "Coupler_Attached",
      "Stopped",
      "EOF"
    ),
    col_types = "nccnnnncccc"
  )
  x$file <- i
  hobo_df <- bind_rows(hobo_df, x)
  rm(x)
}

## clean up the dataframe
hobo_df <- hobo_df %>%
  mutate(
    ## regex extracts site number from file path
    Site = str_extract(file, "\\d{1,6}"),
    
    ## convert date and time columns to date/time format
    dt = paste(Date, Time),
    Date_Time = as.POSIXct(paste(Date, Time),
                           tz = "Etc/GMT-6",
                           format = "%m/%d/%y %I:%M:%S %p")) %>%
  mutate(Site = as.factor(Site)) %>%
  ## select the columns we need to keep
  dplyr::select(Abs_Pres, Temp, Water_Level, Site, Date_Time) %>%
  ## filter rows without water_level
  dplyr::filter(!is.na(Water_Level))
  
## attach units to our columns
units(hobo_df$Water_Level) <- as_units("ft")
units(hobo_df$Temp) <- as_units("°F")
units(hobo_df$Abs_Pres) <- as_units("psi")

  
hobo_df

```

### IQ Plus
```{r iqimport}
#read_csv(here::here("Data/IQPlus/16397-2020_12_31.csv"))

## make a list of files to import
file_paths <- paste0(here("Data/IQPlus"),
                     "/",
                     list.files(path = here("Data/IQPlus"),
                                pattern = ".csv"))

## create a blank tibble to fill
iqplus_df <- tibble()

## loop through file paths to read each file
for (i in file_paths) {
  x <- read_csv(
    i,
    col_types = "nc______n__n______________________nn______n_________"
  )
  x$file <- i
  iqplus_df <- bind_rows(iqplus_df, x)
  rm(x)
}


iqplus_df <- iqplus_df %>%
    mutate(
    ## regex extracts site number from file path
    Site = str_extract(file, "\\d{1,6}")) %>%
  ## use `dplyr::` to specify which rename function to use, just in case
  dplyr::rename(Sample_Number =`Sample number`,
                Date_Time = `Sample time`,
                Depth = `Depth (ft)`,
                Flow = `Flow (ft³/s)`,
                System_In_Water = `System in water (%)`,
                System_Status = `System status (status codes)`,
                Index_Velocity = `Velocity (mean) (ft/s)`) %>%
  dplyr::select(-c(Sample_Number, file)) %>%
  mutate(Date_Time = as.POSIXct(Date_Time,
                                tz = "Etc/GMT-6",
                                format = "%Y-%m-%d %H:%M:%S"))


## attach units to our columns
units(iqplus_df$Depth) <- as_units("ft")
units(iqplus_df$Flow) <- as_units("ft^3/s")
units(iqplus_df$Index_Velocity) <- as_units("ft/s")

iqplus_df
```



## Data Explore

### Pressure Transducers

Plot the depth date over time for all three sites

```{r}
## note, that if the scale name has a space, there is a bug in ggforce
## need to set units_options(parse = FALSE) per:
## https://github.com/thomasp85/ggforce/issues/197

ggplot(hobo_df) +
  geom_line(aes(Date_Time, Water_Level, color = Site)) +
  scale_y_unit(name = "Level", unit = "ft") +
  scale_x_datetime(name = "Date/Time") +
  labs(title = "15-Minute Measured Height") +
  theme_ms()
```

## IQPlus

Plot the depth data drom the IQPlus from all three sites.

```{r}
iqplus_df %>%
  filter(System_Status == 0,
         System_In_Water == 100) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth, color = Site)) +
  facet_wrap(~Site, scales = "free") +
  labs(title = "Stage") +
  theme_ms()
```


Plot the depth-discharge from all three sites as measured by the IQPlus.

```{r}
iqplus_df %>%
  filter(System_Status == 0,
         System_In_Water == 100,
         as.numeric(Flow) > 0) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Site)) +
  scale_y_unit(name = "Flow", unit = "ft^3/s") +
  facet_wrap(~Site, scales = "free") +
  labs(title = "Stage Discharge") +
  theme_ms()
```


```{r}
iqplus_df %>%
  filter(System_Status == 0,
         System_In_Water == 100) %>%
  mutate(Month = lubridate::month(Date_Time, label = TRUE)) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Month), alpha = 0.5) +
  scale_y_unit(name = "Flow", unit = "ft^3/s") +
  facet_wrap(~Site, scales = "free") +
  labs(title = "Stage Discharge") +
  theme_ms()
```

Site `16396` looks pretty good. We still have some data cleaning and sorting on the remaining two sites.



Exploring `16397` further...

```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         #as.numeric(Depth) >= 0.75,
         as.numeric(Flow) > 0) %>%
  mutate(Month = lubridate::month(Date_Time, label = TRUE)) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Month), alpha = 0.5) +
  scale_y_unit(name = "Flow", unit = "ft^3/s") +
  facet_wrap(~Month) +
  labs(title = "Stage Discharge") +
  theme_ms() +
  theme(panel.spacing = unit(2, "points"))
```






Exploring `16882` further...

```{r}
iqplus_df %>%
  filter(Site == "16882",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Flow) > 0) %>%
  mutate(Month = lubridate::month(Date_Time, label = TRUE)) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Month), alpha = 0.5) +
  scale_y_unit(name = "Flow", unit = "ft^3/s") +
  facet_wrap(~Month) +
  labs(title = "Stage Discharge") +
  theme_ms() +
  theme(panel.spacing = unit(2, "points"))
```



The looped rating curve in May and Dec might be a flood flow event. Discharge is typically higher on the rising stage than the falling stage.




```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time <= as.POSIXct("2020-07-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time <= as.POSIXct("2020-07-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time <= as.POSIXct("2020-07-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```



```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-07-01") & 
           Date_Time < as.POSIXct("2020-09-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-07-01") & 
           Date_Time < as.POSIXct("2020-09-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-07-01") & 
           Date_Time < as.POSIXct("2020-09-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```


```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-09-01") & 
           Date_Time < as.POSIXct("2020-10-15")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-09-01") & 
           Date_Time < as.POSIXct("2020-10-15")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-09-01") & 
           Date_Time < as.POSIXct("2020-10-15")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```



```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-10-15") & 
           Date_Time < as.POSIXct("2020-11-25")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-10-15") & 
           Date_Time < as.POSIXct("2020-11-25")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-10-15") & 
           Date_Time < as.POSIXct("2020-11-25")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```


```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-11-25") & 
           Date_Time < as.POSIXct("2020-12-15")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-11-25") & 
           Date_Time < as.POSIXct("2020-12-15")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-11-25") & 
           Date_Time < as.POSIXct("2020-12-15")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```



```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-12-15") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-12-15") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-12-15") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```


```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))
```



##16396

```{r}
iqplus_df %>%
  filter(Site == "16396",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
    filter(Date_Time > as.POSIXct("2020-12-01") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16396",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
    filter(Date_Time > as.POSIXct("2020-12-01") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16396",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
   filter(Date_Time > as.POSIXct("2020-12-01") & 
           Date_Time < as.POSIXct("2020-12-31")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Date_Time)) +
  scale_y_unit(trans = "log10")
```

```{r}
iqplus_df %>%
  filter(Site == "16396",
         #System_Status == 0,
         #System_In_Water == 100,
         #as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Flow))


iqplus_df %>%
  filter(Site == "16396",
         #System_Status == 0,
         #System_In_Water == 100,
         #as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Date_Time, Depth))

iqplus_df %>%
  filter(Site == "16396",
         #System_Status == 0,
         #System_In_Water == 100,
         #as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
  filter(Date_Time > as.POSIXct("2021-01-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow, color = Date_Time))
```

The Jones Formula can be used to estimate the hysteresis or unstready flow depicted in the rating curve. A reduced formula described by @petersen-overleir_modelling_2006 and @zakwan_spreadsheet-based_2018 is used here.

$$
Q = K(h-a)^n\times\sqrt{1 + x\frac{\partial h}{\partial t}}
$$

- $Q$ is discharge

- $h$ is gage height

- $\frac{\partial h}{\partial t}}$ is the partial first order derivative approximated using finite differences. This can be considered the slope or instantaneous rate of change for the function between gage height and time which is estimated using measured stream height values.


Solve for K, a, n, x by minimizing sum of square error (SSE).

$$
SSE = \sum\limits_{i=1}^N[X-Y]^2
$$



In order to incorporate the first derivative of stream height function we will need to (1) split the data set into each sampling period; (2) calculate the derivatives in each dataset; (3) remove rows with NA (basically first record in each sampling event); (4) combine desired datasets based on shape or sample the datasets; (4) refit the function above to one or more datasets.



```{r}
iqplus_df %>%
  filter(Site == "16396",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) >= 0) %>%
  arrange(Date_Time) %>%
  mutate(time_lag = lag(Date_Time, default = Date_Time[1]),
         diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours"))) %>%
  group_split(cumsum(diff_time > 8)) %>%
  ## remove events where max flow did not go over 10 cfs
  keep(~ max(as.numeric(.x$Flow)) > 10) %>%
  map(~select(.x, Date_Time, Depth, Flow)) %>%
  map(~mutate(.x,
              time_lag = lag(Date_Time, default = Date_Time[1]),
              diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours")),
              diff_depth = c(0, diff(.x$Depth)))) %>%
  imap(~mutate(.x, event = as.character(.y))) %>%
  bind_rows() %>%
  filter(!is.na(diff_depth)) %>%
  mutate(J = as.numeric(diff_depth)/as.numeric(diff_time)) -> df_16396



```


```{r}
exponent <- function(x, pow) {
  (abs(x)^pow)*sign(x)
}
SSE <- function(pars, data) {
  Depth = as.numeric(data$Depth)
  J = data$J
  K = pars[1]
  a = pars[2]
  n = pars[3]
  x = pars[4]

  
  preds <- (K*exponent(x = Depth - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))
  #print(preds)
  Q = as.numeric(data$Flow)
  
  ## minimize the sum of square errors per the paper
  sse <- sum((Q - preds)^2, na.rm = TRUE)
  #print(sse)
  sse
}

par <- c(K = 5,
         a = 5,
         n = 2,
         x = 2000)

## limits to the parameter space
lower <- c(0.1, 0.1, 0.1, 0.1)
upper <- c(200, Inf, 10, 5000)

optim.par <- optim(par = par,
                   fn = SSE,
                   data = df_16396,
                   lower = lower,
                   upper = upper,
                   method = "L-BFGS-B")


K <- optim.par$par[[1]]
a <- optim.par$par[[2]]
n <- optim.par$par[[3]]
x <- optim.par$par[[4]]


```



```{r}
df_16396 %>%
  mutate(predicted = (K*exponent(x = as.numeric(Depth) - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))) %>%
  ggplot() +
  geom_point(aes(as.numeric(Depth), as.numeric(Flow), color = "measured"), alpha = 0.5) +
  geom_point(aes(as.numeric(Depth), as.numeric(predicted), color = "predicted"), alpha = 0.2) +
  scale_y_log10() +
  theme_ms()

df_16396 %>%
  mutate(predicted = (K*exponent(x = as.numeric(Depth) - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))) %>%
  ggplot() +
  geom_point(aes(as.numeric(Flow), as.numeric(predicted), color = "measured"), alpha = 0.5) +
  geom_abline(slope = 1) +
  scale_y_log10() + scale_x_log10() +
  theme_ms()
```

```{r}
df_16396 %>%
  mutate(predicted = (K*exponent(x = as.numeric(Depth) - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))) -> df_16396

df_16396_results <- tibble(Site = "16396",
       K = K,
       a = a,
       n = n,
       x = x,
       NSE = hydroGOF::NSE(df_16396$predicted, as.numeric(df_16396$Flow)),
       RMSE = hydroGOF::rmse(df_16396$predicted, as.numeric(df_16396$Flow)))
df_16396_results
```


## Site 16397

This site was pooled from August through November 2020. It appears we can use the standard power function [@venetis1970note]:
$$
Q = K(H-H_0)^z
$$



```{r}
iqplus_df %>%
  filter(Site == "16397",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 1, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  filter(Date_Time > as.POSIXct("2020-11-01")) %>%
  arrange(Date_Time) %>%
  mutate(time_lag = lag(Date_Time, default = Date_Time[1]),
         diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours"))) %>%
  group_split(cumsum(diff_time > 8)) %>%
  ## remove events where max flow did not go over 10 cfs
  keep(~ max(as.numeric(.x$Flow)) > 10) %>%
  map(~select(.x, Date_Time, Depth, Flow)) %>%
  map(~mutate(.x,
              time_lag = lag(Date_Time, default = Date_Time[1]),
              diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours")),
              diff_depth = c(0, diff(.x$Depth)))) %>%
  imap(~mutate(.x, event = as.character(.y))) %>%
  bind_rows() %>%
  filter(!is.na(diff_depth)) %>%
  mutate(J = as.numeric(diff_depth)/as.numeric(diff_time)) -> df_16397



SSE <- function(pars, data) {
  Depth = as.numeric(data$Depth)
  K = pars[1]
  H_0 = pars[2]
  Z = pars[3]

  
  preds <- K*(as.numeric(Depth) - H_0)^Z
  #print(preds)
  Q = as.numeric(data$Flow)
  
  ## minimize the sum of square errors per the paper
  sse <- sum((Q - preds)^2, na.rm = TRUE)
  #print(sse)
  sse
}

par <- c(K = 1,
         H_0 = .001,
         Z = 2)

## limits to the parameter space
lower <- c(0.001, 0, 0.1)
upper <- c(Inf, 0.26, Inf)

optim.par <- optim(par = par,
                   fn = SSE,
                   data = df_16397,
                   lower = lower,
                   upper = upper,
                   method = "L-BFGS-B")

optim.par

K <- optim.par$par[[1]]
H_0 <- optim.par$par[[2]]
Z <- optim.par$par[[3]]

```


```{r}
df_16397 %>%
  mutate(predicted = K*(as.numeric(Depth) - H_0)^Z) %>%
  ggplot() +
  geom_point(aes(as.numeric(Depth), as.numeric(Flow), color = "measured"), alpha = 0.5) +
  geom_point(aes(as.numeric(Depth), as.numeric(predicted), color = "predicted"), alpha = 0.2) +
  #scale_y_log10() +
  theme_ms()

df_16397 %>%
  mutate(predicted = K*(as.numeric(Depth) - H_0)^Z) %>%
  ggplot() +
  geom_point(aes(as.numeric(Flow), as.numeric(predicted), color = "measured"), alpha = 0.5) +
  geom_abline(slope = 1) +
  #scale_y_log10() + scale_x_log10() +
  theme_ms()
```


```{r}
df_16397 %>%
  mutate(predicted = K*(as.numeric(Depth) - H_0)^Z) -> df_16397

df_16397_results <- tibble(Site = "16397",
       K = K,
       H_0 = H_0,
       Z = Z,
       NSE = hydroGOF::NSE(df_16397$predicted, as.numeric(df_16397$Flow)),
       RMSE = hydroGOF::rmse(df_16397$predicted, as.numeric(df_16397$Flow)))
df_16397_results
```


## site 16882

```{r}
iqplus_df %>%
  filter(Site == "16882",
         System_Status == 0,
         System_In_Water == 100,
         as.numeric(Depth) >= 0.26, ## minimum operating depth
         as.numeric(Flow) > 0) %>%
  ## remove outliers identified in box plots under 1 foot depth
  filter(as.numeric(Depth) < 1 & as.numeric(Flow) < 8 |
           as.numeric(Depth) > 1) %>% 
  arrange(Date_Time) %>%
  mutate(time_lag = lag(Date_Time, default = Date_Time[1]),
         diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours"))) %>%
  group_split(cumsum(diff_time > 8)) %>%
  ## remove events where max flow did not go over 10 cfs
  keep(~ max(as.numeric(.x$Flow)) > 10) %>%
  map(~select(.x, Date_Time, Depth, Flow)) %>%
  map(~mutate(.x,
              time_lag = lag(Date_Time, default = Date_Time[1]),
              diff_time = as.numeric(difftime(Date_Time, time_lag, units = "hours")),
              diff_depth = c(0, diff(.x$Depth)))) %>%
  imap(~mutate(.x, event = as.character(.y))) %>%
  bind_rows() %>%
  filter(!is.na(diff_depth)) %>%
  mutate(J = as.numeric(diff_depth)/as.numeric(diff_time)) -> df_16882

df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-11-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))

df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-05-01") &
           Date_Time < as.POSIXct("2020-06-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))


df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-10-01") &
           Date_Time < as.POSIXct("2020-11-01")) %>%
  ggplot() +
  geom_point(aes(Depth, Flow))




```


```{r}

SSE <- function(pars, data) {
  Depth = as.numeric(data$Depth)
  J = data$J
  K = pars[1]
  a = pars[2]
  n = pars[3]
  x = pars[4]

  
  preds <- (K*exponent(x = Depth - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))
  #print(preds)
  Q = as.numeric(data$Flow)
  
  ## minimize the sum of square errors per the paper
  sse <- sum((Q - preds)^2, na.rm = TRUE)
  #print(sse)
  sse
}

par <- c(K = 5,
         a = 5,
         n = 2,
         x = 2000)

## limits to the parameter space
lower <- c(0.1, 0.1, 0.1, 0.1)
upper <- c(200, Inf, 10, 5000)

optim.par <- optim(par = par,
                   fn = SSE,
                   data = df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-11-01") & !is.infinite(J)),
                   lower = lower,
                   upper = upper,
                   method = "L-BFGS-B")


K <- optim.par$par[[1]]
a <- optim.par$par[[2]]
n <- optim.par$par[[3]]
x <- optim.par$par[[4]]


```

```{r}
df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-11-01") & !is.infinite(J)) %>%
  mutate(predicted = (K*exponent(x = as.numeric(Depth) - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))) %>%
  ggplot() +
  geom_point(aes(as.numeric(Depth), as.numeric(Flow), color = "measured"), alpha = 0.5) +
  geom_point(aes(as.numeric(Depth), as.numeric(predicted), color = "predicted"), alpha = 0.2) +
  #scale_y_log10() +
  theme_ms()

df_16882 %>%
  filter(Date_Time > as.POSIXct("2020-11-01") & !is.infinite(J)) %>%
  mutate(predicted = (K*exponent(x = as.numeric(Depth) - a, pow = n)) * exponent(x = (1 + x * J), pow = (1/2))) %>%
  ggplot() +
  geom_point(aes(as.numeric(Flow), as.numeric(predicted), color = "measured"), alpha = 0.5) +
  geom_abline(slope = 1) +
  #scale_y_log10() + scale_x_log10() +
  theme_ms()
```

